---
title: "Religious Locations of Toronto TBD"
author: 
  - Eunkyo Oh
thanks: "Code and data are available at: https://github.com/oheunkyo/paper1."
date: "`r format(Sys.time(), '%d %B %Y')`"
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
output:
  bookdown::pdf_document2
toc: FALSE
bibliography: references.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
```

# Introduction

Use the first setnece of the abstract to make a paragraoh

Second paragraph is abou twhat was done and what was found. (setence 2,3 of abstaact)

Third paragph is about  implications (final sentence)

Final paragraph about the remainder of the paper. Section something explains the data section something covbers results....


You can and should cross-reference sections and sub-sections. For instance, Section \@ref(data). R Markdown automatically makes the sections lower case and adds a dash to spaces to generate labels, for instance, Section \@ref(first-discussion-point). 
\@ref(additional-details)


# Data

Paragraph of two describing data broadly


Show an extract of the dataset wiht knitr and cable like how it is on the book (Table \@ref(tab:dataextract)) 

```{r dataextract}
library(tidyverse)
library(knitr)
religious_locations <- read_csv(here::here("inputs/data/religious_locations.csv"))

religious_locations |> 
  select(MUNICIPALITY, FTH_ORGANIZATION, FTH_FAITH) |>
  slice(1:10) |> 
  kable(
    caption = "First ten rows of a dataset of places of worship for religious
    purposes for in Greater Toronto Area",
    col.names = c("Municipality", "Name of Location", "Religion of Practice"),
    digits = 1,
    booktabs = TRUE, 
    linesep = ""
  )
```


Also bills and their average (Figure \@ref(fig:billssssss)). (Notice how you can change the height and width so they don't take the whole page?)

```{r billssssss, fig.cap="More bills of penguins", echo = FALSE, fig.width=8, fig.height=4}

# This needs to be about the random data tha tI sasmpled?
#ggplot(penguins, aes(x = island, fill = species)) +
#  geom_bar(alpha = 0.8) +
#  scale_fill_manual(values = c("darkorange","purple","cyan4"),
#                    guide = FALSE) +
#  theme_minimal() +
#  facet_wrap(~species, ncol = 1) +
#  coord_flip()
```






# Model

\begin{equation}
Pr(\theta | y) = \frac{Pr(y | \theta) Pr(\theta)}{Pr(y)}  (\#eq:bayes)
\end{equation}

Equation \@ref(eq:bayes) seems useful, eh?

Here's a dumb example of how to use some references: In paper we run our analysis in `R` [@citeR]. We also use the `tidyverse` which was written by @thereferencecanbewhatever If we were interested in baseball data then @citeLahman could be useful. 

We can use maths by including latex between dollar signs, for instance $\theta$.


# Results

# Discussion

## First discussion point

If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this. 

## Second discussion point

## Third discussion point

## Weaknesses and next steps

Weaknesses and next steps should also be included.

\newpage

\appendix

# Appendix {-}


# Additional details


\newpage


# References


